---
title: Data Visualization
output: html_document
---

```{r, include=FALSE}
library(ggplot2)
```

```{r, eval=FALSE}
ggplot2::ggplot(data = DATA) +
	geom_function(mapping = MAPPINGS)
```

### Exercise 3.2.4 - Introduction
1. Run `ggplot(data = mpg)`. What do you see?  
```{r}
ggplot(data = mpg)
```

2. How many rows are in `mpg`? How many columns?  
```{r}
nrow(mpg)

ncol(mpg)
```

3. What does the `drv` variable describe? Read the help for `?mpg` to find out  
```{r}
?mpg
```

__Answer:__ the type of drive train, where f = front-wheel drive, r = rear wheel
drive, 4 = 4wd  

4. Make a scatterplot of `hwy` vs `cyl`.  
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(y = hwy, x = cyl))
```

5. What happens if you make a scatterplot of `class` vs `drv`? 
Why is the plot not useful?  
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = class, y = drv))
```

### Exercise 3.3.1 - Aesthetic mappings
1. What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

2. Which variables in mpg are categorical? 
Which variables are continuous? 
(Hint: type ?mpg to read the documentation for the dataset). 
How can you see this information when you run mpg?  
```{r}
str(ggplot2::mpg)
```

3. Map a continuous variable to `color`, `size`, and `shape`. 
How do these aesthetics behave differently for categorical vs. 
continuous variables?
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty, color = hwy))

ggplot(data = mpg) + 
       	geom_point(mapping = aes(x = displ, y = cty, color = class))
```

```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty, size = hwy))

ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty, size = class))
```

```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y =cty), shape = 22)

ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y =cty, shape = class))

ggplot(data = mpg) + 
	geom_point(mapping = aes(x = displ, y = cty, shape = hwy))
```

4. What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty, 
			     color = drv, 
			     shape = drv))
```

5. What does the stroke aesthetic do? 
What shapes does it work with? (Hint: use ?geom_point)  
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty, stroke = 1),
		 shape = 1)

ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty, stroke = 5),
		 shape = 1)
```

6. What happens if you map an aesthetic to something other than a variable name,
like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty,
			     color = displ < 5))
```

### Exercise 3.5.1 - Facets
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty)) +
	facet_wrap(~ class, nrow = 2)

ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty)) +
	facet_grid(drv ~ cyl)

ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty)) +
	facet_grid(.~ cyl)
```

1. What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty)) +
	facet_wrap(~ hwy)
```

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? 
How do they relate to this plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

3. What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

4. What are the advantages to using faceting instead of the colour aesthetic? 
What are the disadvantages? 
How might the balance change if you had a larger dataset?
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = hwy, color = class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

5. Read `?facet_wrap`.  
What does `nrow` do? Number of rows.  
What does `ncol` do? Number of columns.  
What other options control the layout of the individual panels? 
`as.table`, `witch`, `drop`, `dir`, `strip.position`.  
Why doesn’t `facet_grid()` have nrow and ncol arguments? 
Because the combination of variabe A and B are fixed.  

6. When using `facet_grid()` you should usually put the variable with more 
unique levels in the columns. Why?  
```{r}
ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty)) +
	facet_grid(drv ~ cyl)

ggplot(data = mpg) +
	geom_point(mapping = aes(x = displ, y = cty)) +
	facet_grid(cyl ~ drv)

```

### Exercise 3.6.1 - Geometric objects
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

```

1. What geom would you use to draw a line chart? `geom_line()`  
A boxplot? `geom_boxplot()`  
A histogram? `geom_histogram()`  
An area chart? `geom_area()`  

2. Run this code in your head and predict what the output will look like. 
Then, run the code in R and check your predictions.  
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3. What does `show.legend = FALSE` do? Omit the legend.  
What happens if you remove it? The legend is not displayed.  
Why do you think I used it earlier in the chapter? To compare 3 plots.  

4. What does the `se` argument to `geom_smooth()` do? 
Display the confidence interval around smooth.   

5. Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

6. Recreate the R code necessary to generate the following graphs.  
```{r}
fig = ggplot(data = mpg,
	   mapping = aes(x = displ, y = hwy))

fig + geom_point() +
	geom_smooth(se = FALSE)

fig + geom_point() + 
	geom_smooth(mapping = aes(group = drv),
			      se = FALSE)

fig + geom_point(mapping = aes(color = drv)) +
	geom_smooth(mapping = aes(color = drv),
		  se = FALSE)

fig + geom_point(mapping = aes(color = drv)) +
	geom_smooth(mapping = aes(x = displ, y = hwy),
		  se = FALSE)

fig + geom_point(mapping = aes(color = drv)) + 
	geom_smooth(mapping = aes(linetype = drv),
		  se = FALSE)

fig + geom_point(mapping = aes(stroke = 3),
	       shape = 1,
	       color = "white") +
	geom_point(mapping = aes(color = drv))
	
```

### Exercise 3.7.1 - Statistical transformation
1. What is the default geom associated with stat_summary()? `pointrange`  
How could you rewrite the previous plot to use that geom function instead of the 
stat function?
```{r}
ggplot(data = diamonds) +
	geom_pointrange(mapping = aes(x = cut, y = depth),
		      stat = "summary",
		      fun.min = min,
		      fun.max = max,
		      fun = median)
```

2. What does `geom_col()` do? 
Display a bar chart. The heights of the bars represent values in the data.  
How is it different to `geom_bar()`? 
`geom_bar()` makes the height of the bar proportional to the number of cases in 
each group, e.g. count.  

3. Most `geoms` and `stats` come in pairs that are almost always used in concert.
Read through the documentation and make a list of all the pairs. 
What do they have in common?  
The solution is [here](https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html#exercise-3.7.3)  

4. What variables does `stat_smooth()` compute? 
What parameters control its behaviour?  
* y or x: predicted value  
* ymin or xmin: lower pointwise confidence interval around the mean  
* ymax or xmax: upper pointwise confidence interval around the mean  
* se: standard error  
See [here](https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html#exercise-3.7.4)  

5. In our proportion bar chart, we need to set group = 1. 
Why? In other words what is the problem with these two graphs?  
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop)))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
```

### Exercise 3.8.1 - Position adjustments
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

1. What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

### Exercise 3.9.1 - Coordinate system
1. Turn a stacked bar chart into a pie chart using `coord_polar()`.
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity)) +
	coord_polar()
```

2. What does `labs()` do? Read the documentation.

3. What’s the difference between `coord_quickmap()` and `coord_map()`?  
Solution [here](https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html#exercise-3.9.3)  

4. What does the plot below tell you about the relationship between city 
and highway mpg?  
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

Why is `coord_fixed()` important? 
What does `geom_abline()` do? Plot the line.  


