---
title: "Data transformation"
output: html_document
---

```{r}
library(tidyverse)
library(nycflights13)
```

```{r}
flights
```

## Exercise 5.2 --> dplyr::filter()
1. Find all flights that..
```{r}
# Had an arrival delay of two or more hours
flights %>% filter(arr_delay >= 120)
```

```{r}
# Flew to Houston (IAH or HOU)
houston1 <- flights %>% filter(dest == "IAH" | dest == "HOU")

houston2 <- flights %>% filter(dest %in% c("IAH", "HOU"))

identical(houston1, houston2)

houston1
```

```{r}
# Were operated by United, American, or Delta
airlines

carrier1 <- flights %>% filter(carrier == "UA" | carrier == "AA" | carrier == "DL")

carrier2 <- flights %>% filter(carrier %in% c("UA", "AA", "DL"))

identical(carrier1, carrier2)

carrier1
```

```{r}
# Departed in summer (July, August, and September)
season1 <- flights %>% filter(month == 7 | month == 8 | month == 9)

season2 <- flights %>% filter(month %in% c(7, 8, 9))
season3 <- flights %>% filter(month %in% 7:9)

season4 <- flights %>% filter(between(month, 7, 9))


identical(season1, season2)
identical(season2, season3)
identical(season3, season4)

season1
```

```{r}
# Arrived more than two hours late, but didn’t leave late
flights %>% filter(arr_delay > 120, dep_delay <= 0)
```

```{r}
# Were delayed by at least an hour, but made up over 30 minutes in flight
flights %>% filter(dep_delay >= 60, dep_delay - arr_delay > 30)
```

```{r}
# Departed between midnight and 6am (inclusive)
summary(flights$dep_time)

flights %>% filter(dep_time == 2400 | dep_time <= 600)

```

3.
```{r}
# How many flights have a missing dep_time? 
flights %>% filter(is.na(dep_time) == TRUE) %>% count()

# What other variables are missing? What might these rows represent?
```

## Exercise 5.3 - dplyr::arrange()
```{r}
# 1. sort all missing values to the start
flights %>% arrange(desc(is.na(dep_time)))
```

```{r}
# 2. Sort flights to find the most delayed flights
flights %>% arrange(desc(dep_delay))

# Find the flights that left earliest
flights %>% arrange((dep_delay))
```

```{r}
# 3. Find the fastest (highest speed) flights
flights %>% arrange(desc(distance / air_time))

flights %>% mutate(speed = distance / air_time) %>% arrange(desc(speed))

```

```{r}
# 4. Which flights travelled the farthest? 
flights %>% arrange(desc(distance))

# Which travelled the shortest?
flights %>% arrange(distance)
```

## Exercise 5.4. - dplyr::select()
1. select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`
```{r}
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)

flights %>% select(starts_with(c("dep_", "arr_")))
```

2. What happens if you include the name of a variable multiple times in a 
`select()` call?
```{r}
flights %>% select(dep_time, dep_delay, dep_time, dep_time)
```

3. What does the `any_of()` function do?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

flights %>% select(any_of(vars))

flights %>% select(vars)

```

4. Does the result of running the following code surprise you?
```{r}
select(flights, contains("TIME"))

select(flights, contains("TIME", ignore.case = FALSE))
```

## Exercise 5.5. - dplyr::mutate()
1. Convert `dep_time` and `sched_dep_time` to number of minutes since midnight  
```{r}
convertMin = function(x) {(x %/% 100 * 60 + x %% 100) %% 1440}

flights %>% mutate(dep_time_min = convertMin(dep_time),
	         sched_dep_time_min = convertMin(sched_dep_time)
	         )
```

2. Compare `air_time` with `arr_time - dep_time`. 
What do you expect to see? What do you see? What do you need to do to fix it?  
```{r}
flights %>% mutate(flight_time = arr_time - dep_time,
	         compare = air_time - flight_time) %>%
	select(air_time, flight_time, compare)

```

I expect to see `air_time` == `arr_time - dep_time`.  
I see the problem:  
* `air_time` unit is minute  
* `arr_time - dep_time` unit is in format HMM or HHMM  
=> Solution:  
* convert `arr_time` and `dep_time` to minutes  
* compare
```{r}
flights %>% mutate(dep_time_min = convertMin(dep_time),
	         arr_time_min = convertMin(arr_time),
	         flight_time = arr_time_min - dep_time_min,
	         compare = air_time - flight_time
	         ) %>%
	select(air_time, flight_time, compare,
	       dep_time, dep_time_min, 
	       arr_time, arr_time_min 
	       )
```

3. Compare `dep_time`, `sched_dep_time`, and `dep_delay`. 
How would you expect those three numbers to be related?  
```{r}
flights %>% mutate(dep_time_min = convertMin(dep_time),
	         sched_dep_time_min = convertMin(sched_dep_time),
	         dep_delay_expect = dep_time_min - sched_dep_time_min,
	         compare = dep_delay - dep_delay_expect) %>%
	select(dep_time, dep_time_min,
	       sched_dep_time, sched_dep_time_min,
	       dep_delay_expect,
	       dep_delay,
	       compare)
```

4. Find the 10 most delayed flights using a ranking function. 
How do you want to handle ties? Carefully read the documentation for `min_rank()` 

5. What does 1:3 + 1:10 return? Why?  
```{r}
1:3 + 1:10
```

6. What trigonometric functions does R provide?  

## Exercise 5.6. - dplyr::summarise()


## Exercise 5.7 - Grouped mutates (and filters)



